Coherent optical association of a single molecule

### 2020-10-27 paper.tex
Update on Overleaf.

\begin{abstract}
  We report on coherent association of a single weakly-bound NaCs molecule in an optical tweezer
  through an optical Raman transition without the use of a Feshbach resonance. Our scheme borrows transition dipole moment while reducing photon scattering by selecting a deeply bound electronic excited intermediate state.
  Starting from two atoms in their relative motional ground state,
  we achieve optical transfer efficiency of $69\%$.
  The molecule has a binding energy of $770.200516(24) \mathrm{MHz}$ at $8.8~\mathrm{G}$
  with more than $60\%$ of the molecule created in the motional ground state.
  This technique is general without relying on narrow excited state lines or Feshbach resonances and could allow a wider range of molecular species to be assembled atom-by-atom.
\end{abstract}

@@

Such ultracold molecular association takes advantage of the much developed cooling and trapping techniques for atoms as a starting point. To overcome the challenges of small wavefunction overlap and the large release of binding energy of converting atoms to deeply-bound molecules, a  two-step approach has been established to first associate atom pairs into weakly-bound molecules, and then transfer the molecules from this single internal state to a desired rovibrational and electronic state~\cite{Danzl2008, Ni2008,Lang2008, Takekoshi2014, Molony2014, Park2015, Guo2016, Kondov2019, Voges2020}.
So far, all of such association processes utilized a magnetic Feshbach scattering resonance and have been applied to bialkali molecules. The only exceptions are Sr$_2$ where narrow linewidth excited states are available and optical association can be driven coherently~\cite{Reinaudi2012,Stellmer2012} and $^{87}$Rb$^{85}$Rb where weakly bound molecules of a binding energy of a couple MHz exists~\cite{He331}. These requirements  limit the generality of previous association techniques.
% The requirement of a Feshbach resonance or a MHz-level binding energy bound state to enhance atom-to-molecule wavefunction overlap or the existence of narrow excited state lines limits the generality of previous association techniques. % to more diverse species.

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule using a two-photon optical Raman transfer via an electronic excited state, schematically shown in Fig.~\ref{f-theory}A, neither using a Feshbach resonance, few MHz-level bound states, nor a narrow excited state. The resulting single molecule is in a single internal quantum state and predominately in its motional ground state.  Our scheme is based on a choice of vibrational state of the electronic excited state $c^3\Sigma^+(\Omega = 1)$ that has the theoretical best Raman Rabi frequency to photon scattering ratio. To further increase this ratio and reduce technical requirements such as intensity stability, we choose an initial and final state where the matrix elements with the excited state are as balanced as possible. This approach minimizes the reliance on system specific properties and could therefore be applied to creating other molecular species or larger molecules atom-by-atom with full quantum state control.
% (also this result represents full quantum-state control of single [neutral] molecules, the second time ever (first being our NaCs Feshbach molecules))

@@

The essence of an optical Raman transfer can be illustrated using a three-level system (Fig.~\ref{f-theory}A), where the initial atomic state, the target weakly-bound molecular state are coupled to an intermediate state by two photons, $\Omega_a$ and $\Omega_m$, with one-photon detuning $ \Delta $, and two-photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $c^3\Sigma^+$)
The transfer Raman Rabi Rate, $\Omega_R=\Omega_a\Omega_m / 2\Delta$, is accompanied by a photon scattering rate $\Gamma_s$ that is a sum of all coupling sources%,
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
~\cite{Wineland2003}.
Unlike for Raman transitions in atoms, the two coupling matrix elements here are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and therefore the scattering is predominantly from the target molecular state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single photon detuning, $ \Delta $, so the target molecular state can scatter off both beams roughly equally. Thus, the scattering is given by $ \Gamma_s=\Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth \footnote{We choose the two beams to have equal power, which gives the highest Raman Rabi rate at a fixed total power. Thus, this results in a simple factor of 2 coming from scattering off 2 beams.}.
The ratio between the Raman Rabi frequency and the scattering rate is therefore $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

Pioneering experiments used weakly-bound excited states as the intermediate state in the Raman transition to ensure a large Raman Rabi frequency~\cite{Wynar2000,Rom2004}. However, for a complete picture, the many excited vibrational states in a molecular potential as well as the excited atomic continuum need to be considered. The total scattering rate and Raman Rabi rate become a sum of the scattering rates and Raman Rabi rates over all possible intermediate states. With these considerations, using a weakly-bound excited state as the intermeidate state suffers from strong scattering of the nearby excited atomic continuum rendering it incoherent, resulting in molecule loss. This scattering is proportional to $1/\delta_{thresh}^2$, where $\delta_{thresh}$ is the detuning from the dissociation threshold, and thus can  be made smaller by detuning away from the dissociation threshold.
%With the multiple excited vibrational states present in a molecular potential as intermediate states, the total scattering rate and Raman Rabi rate become a sum over the scattering rates and Raman Rabi rates of the individual intermediate states. Furthermore, for a more complete calculation, we also need to account for the effects of the excited atomic continuum, which is particularly important for the target molecular state. The sum of the squares of the wavefunction overlap between the target weakly-bound molecular state and all the excited molecular intermediate states considered is only about 2\%, suggesting that there is significant matrix element between the target molecular state and the excited atomic continuum. (See SM)

>>>>

\begin{abstract}
  We report on coherent association of a single weakly-bound NaCs molecule in an optical tweezer
  through an optical Raman transition without the use of a Feshbach resonance. The Raman scheme uses a deeply bound electronic excited intermediate state to achieve a large transition dipole moment while reducing the photon scattering. 
  Starting from two atoms in their relative motional ground state,
  we achieve an optical transfer efficiency of $69\%$.
  The molecule has a binding energy of $770.200516(24) \mathrm{MHz}$ at $8.8~\mathrm{G}$ (\textcolor{red}{Unresolved comment from Jon: with so many digits on the resonance frequency, we may need to mention more digits on the B field number and/or also state the tweezer power. We can state the tweezer power for where we measure it, but the transmission through the objective, etc. I believe is only estimated.})
  with more than $60\%$ of the molecule created in the motional ground state.
  This technique does not rely on narrow excited state lines or Feshbach resonances and will allow a wider range of molecular species to be assembled atom-by-atom.
\end{abstract}

@@

Molecular association of ultracold atoms takes advantage of the cooling and trapping techniques that have been developed for atoms. Associating atoms into deeply-bound molecules is challenging because of the small wavefunction overlap between the free-atom and molecular state and the release of a large amount of binding energy. A two-step approach has been demonstrated to first associate atom pairs into weakly-bound molecules, and then transfer the molecules from this single internal state to a desired rovibrational and electronic state releasing the binding energy via stimulated emission~\cite{Danzl2008, Ni2008,Lang2008, Takekoshi2014, Molony2014, Park2015, Guo2016, Kondov2019, Voges2020}.
So far, most molecular association has been achieved for bialkali molecules by magnetoassociation using a magnetic Feshbach scattering resonance. The only exceptions are Sr$_2,$ where narrow linewidth ($\sim 20 ~\mathrm{kHz}$) excited states are available and optical association can be driven coherently,~\cite{Reinaudi2012,Stellmer2012} and $^{87}$Rb$^{85}$Rb, where weakly bound molecules with a couple MHz binding energy exists~\cite{He331}. With these requirements, dimers between alkali atoms and non-magnetic atoms, or molecules involving atoms without narrow intercombination lines remain difficult to associate. 
% These requirements limit the generality of previous association techniques.
% The requirement of a Feshbach resonance or a MHz-level binding energy bound state to enhance atom-to-molecule wavefunction overlap or the existence of narrow excited state lines limits the generality of previous association techniques. % to more diverse species.

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule using a two-photon optical Raman transfer via an electronic excited state, schematically shown in Fig.~\ref{f-theory}A. The scheme does not rely on a Feshbach resonance, few MHz level bound states, or a narrow excited state. The resulting single molecule is in a well-defined internal quantum state and predominately in its motional ground state. A vibrational state of the electronic excited state $c^3\Sigma^+(\Omega = 1)$ is used as the intermediate state in the Raman scheme, and is chosen to minimize the theoretical photon scattering during a Raman Rabi oscillation. To further reduce the photon scattering and sensitivity to laser intensity noise, we choose the initial and final state to balance the two Rabi frequencies as much as possible. This system-independent approach can be used for creating new molecules atom-by-atom with full quantum state control. 
% and could therefore be applied to creating other molecular species or larger molecules atom-by-atom with full quantum state control.
% (also this result represents full quantum-state control of single [neutral] molecules, the second time ever (first being our NaCs Feshbach molecules))

@@

The essence of an optical Raman transfer can be illustrated using a three-level system (Fig.~\ref{f-theory}A), where the initial atomic state, and the target weakly-bound molecular state are coupled to an intermediate state by two laser frequencies with Rabi frequencies, $\Omega_a$ and $\Omega_m$, with one-photon detuning $ \Delta $, and two-photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $c^3\Sigma^+$)
The transfer Raman Rabi Rate, $\Omega_R=\Omega_a\Omega_m / 2\Delta$, is accompanied by a photon scattering rate $\Gamma_s$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
~\cite{Wineland2003}.
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and therefore the scattering is predominantly from the target molecular state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single photon detuning, $ \Delta $, so the target molecular state can scatter off both beams roughly equally. Thus, the scattering is given by $ \Gamma_s=\Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth \footnote{We choose the two beams to have equal power, which gives the highest Raman Rabi rate at a fixed total power. Thus, this results in a simple factor of 2 coming from scattering off 2 beams.}.
The ratio between the Raman Rabi frequency and the scattering rate is therefore $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

Early experiments used weakly-bound excited states as the intermediate state in the Raman transition to ensure a large Raman Rabi frequency~\cite{Wynar2000,Rom2004}. However, for a complete picture, the many excited vibrational states in a molecular potential as well as the excited atomic continuum need to be considered. The total scattering rate and Raman Rabi rate become a sum of the scattering rates and Raman Rabi rates over all possible intermediate states. With these considerations, using a weakly-bound excited state as the intermeidate state suffers from strong scattering of the nearby excited atomic continuum rendering it incoherent, resulting in molecule loss. This scattering is proportional to $1/\delta_{thresh}^2$, where $\delta_{thresh}$ is the detuning from the dissociation threshold, and thus can  be made smaller by detuning away from the dissociation threshold.
%With the multiple excited vibrational states present in a molecular potential as intermediate states, the total scattering rate and Raman Rabi rate become a sum over the scattering rates and Raman Rabi rates of the individual intermediate states. Furthermore, for a more complete calculation, we also need to account for the effects of the excited atomic continuum, which is particularly important for the target molecular state. The sum of the squares of the wavefunction overlap between the target weakly-bound molecular state and all the excited molecular intermediate states considered is only about 2\%, suggesting that there is significant matrix element between the target molecular state and the excited atomic continuum. (See SM)
